---
title: "CCA on Halper & Yang"
author: "Nick Wawee"
date: "August 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/waweenic.BME011/Desktop/Research/Zonation_Yang")
rm(list = ls())
library(ggplot2)
#library(monocle)
library(R.utils)
library(tidyr)
library(plyr)
library(dplyr)
library(Hmisc)
library(xlsx)
#library(forecast)
library(Biobase)
library(Seurat)
library(cowplot)
orderedmatrix<-function(mat,rownumber,colnumber){ # this function will sort a matrix based on the numerical value in the designated row number in ascending order (least to greatest).
  if (colnumber=="NA"){
    valuevec<-as.matrix(mat[rownumber,])
    cellnames<-rownames(valuevec)
    genenames<-rownames(mat)
    valuevec<-as.character(valuevec)
    valuemat<-cbind(cellnames,valuevec)
    numericvec<-sort(as.numeric(valuemat[,2]))
    charactervec<-as.character(numericvec)
    matchingindices<-match(charactervec,valuevec)
    charactermat<-cbind(cellnames[matchingindices],charactervec)
    orderedmatrix<-mat[,matchingindices]
  }else if(rownumber=="NA"){
    valuevec<-as.matrix(mat[,colnumber])
    genenames<-rownames(valuevec)
    cellnames<-colnames(mat)
    valuevec<-as.character(valuevec)
    valuemat<-cbind(genenames,valuevec)
    numericvec<-sort(as.numeric(valuemat[,2]))
    charactervec<-as.character(numericvec)
    matchingindices<-match(charactervec,valuevec)
    orderedmatrix<-mat[matchingindices,]
  }
}
mat2numericmat<-function(mat){# this function converts a matrix of class matrix to a numeric matrix
  if (class(mat)=="matrix"){
    newmat<-matrix(0, length(mat[,1]),length(mat[1,]))
    cnames<-colnames(mat)
    rnames<-rownames(mat)
    for (i in 1:length(newmat[,1])){
      newmat[i,]<-sapply(mat[i,],as.numeric)
    }
    row.names(newmat)<-rnames
    colnames(newmat)<-cnames
    mat2numericmat<-newmat
  }else{stop('the matrix is not of the matrix class. class(mat)!=matrix')}
}
scalefun<-function(evenlyspaceddata,scaling){
  if(scaling=="statistics" & sd(evenlyspaceddata)!=0){
    scalefun<-(evenlyspaceddata - mean(evenlyspaceddata))/sd(evenlyspaceddata)
  } else if (scaling=="0and1"){
    range=max(evenlyspaceddata)-min(evenlyspaceddata)
    if (range!=0){
      scalefun<-(evenlyspaceddata-min(evenlyspaceddata))/(max(evenlyspaceddata)-min(evenlyspaceddata))
    }else{scalefun<-evenlyspaceddata}
  }
}

topminmax<-function(vector, min, max, number, mag){ #takes column vector in matrix format and finds the 50 most minimum or maximum values in it and returns a new matrix vector.
newmat<-matrix(0,number,2)
if (mag=="Yes"){
  vector<-abs(vector)
}
if (max=="NA"){
  for (i in 1:number){
    minval<-min(vector)
    minind<-which.min(vector)
    minname<-rownames(vector)[minind]
    othernames<-rownames(vector)[-minind]
    newmat[i,]<-c(minval,minname)
    vector<-as.matrix(vector[-minind])
    row.names(vector)<-othernames
    }
  }
if (min=="NA"){
    for (i in 1:number){
    maxval<-max(vector)
    maxind<-which.max(vector)
    maxname<-rownames(vector)[maxind]
    othernames<-rownames(vector)[-maxind]
    newmat[i,]<-c(maxval,maxname)
    vector<-as.matrix(vector[-maxind])
    row.names(vector)<-othernames
    }
  }
row.names(newmat)<-newmat[,2]
newmat<-newmat[,-2]
newmat<-sapply(newmat, as.numeric)
newmat<-as.matrix(newmat)
return(newmat)
}
idandname<-function(idnamemat, geneid, genename){#This function returns the desired gene name(s) or id(s) with the input of a matrix that has row names corresponding to each gene identifier and the 1st column contains the respective short name
  if (geneid=="NA"){
    geneid.r=rownames(idnamemat)[match(genename,idnamemat[,1])]
    return(geneid.r)
  }
  if (genename=="NA"){
    genename.r=idnamemat[match(geneid,rownames(idnamemat)),1]
    return(genename.r)
  }
  
}
```

#Loading

```{r}
#Loading
loadmaty<-read.table("Yang_ReadCount_ERCC_Converted.txt", fill = TRUE, stringsAsFactors = FALSE)
loadmaty2<-read.table("pseudotimemat.txt", stringsAsFactors = FALSE)
zonatedmat<-read.table("OrderedPtMatrix_Filtered_1.txt", stringsAsFactors = FALSE)
zonatedmat<-as.data.frame(zonatedmat)

#Feature Data
zgenes<-rownames(zonatedmat)
zonatedmat<-cbind(zgenes,zonatedmat)
zonatedmat$zgenes<-as.character(zonatedmat$zgenes)
zonatedmat2<-zonatedmat %>% 
  separate_rows(zgenes, sep=";")
row.names(zonatedmat2)<-zonatedmat2[,1]

# Testing separte_rows ---------------------------------
testexp<-zonatedmat2[c(which(rownames(zonatedmat2)=="Apr3"),which(rownames(zonatedmat2)=="0610007C21Rik")),]
testexp2<-zonatedmat[which(rownames(zonatedmat)=="0610007C21Rik;Apr3"),]
#View(rbind(testexp,testexp2))
#---------------------------------------------------------------------------------------------------------------------------------------

loadmaty[,2]<-as.character(loadmaty[,2])
matchind<-match(rownames(zonatedmat2),loadmaty[,2])
matchind<-matchind[-which(is.na(matchind))]
fdmat<-as.matrix(loadmaty[matchind,1:2])
row.names(fdmat)<-fdmat[,1]
fdmat<-as.matrix(fdmat[,-1])
colnames(fdmat)<-"gene_short_name"
fd<-as.data.frame(fdmat)
fd<-new("AnnotatedDataFrame", data=fd)

#Pheno Data
#Yang
cellnamesy<-as.matrix(colnames(loadmaty2)[which(loadmaty2[2,]=="Hepatoblast"|loadmaty2[2,]=="Hepatocyte")])
colnames(cellnamesy)<-cellnamesy[1,]
cellnamesy<-cbind(cellnamesy,rep("Embryonic",length(cellnamesy[,1])))
colnames(cellnamesy)<-c("Cell_Name", "Cell_Type")
row.names(cellnamesy)<-cellnamesy[,1]
pt<-loadmaty2[1,match(rownames(cellnamesy),colnames(loadmaty2))]
pt<-as.numeric(pt)
pt<-scalefun(pt, "0and1")
statey<-loadmaty2[2,match(rownames(cellnamesy),colnames(loadmaty2))]
statey<-t(statey)
colnames(statey)<-"State0"
statey2<-statey
colnames(statey2)<-"State2"
cellnamesy<-cbind(cellnamesy,pt,statey,statey2)

#Zonation
cellnamesz<-as.matrix(colnames(zonatedmat2)[2:length(zonatedmat2)])
row.names(cellnamesz)<-cellnamesz
cellnamesz<-cbind(cellnamesz, rep("Zonation",length(cellnamesz)))
colnames(cellnamesz)<-c("Cell_Name", "Cell_Type")
#cellnamesz<-cellnamesz[-length(cellnamesz[,1]),]
cellnamesz<-cbind(cellnamesz,t(zonatedmat2[c(1,4,5),-1]))
cellnamesz[,3]<-scalefun(as.numeric((cellnamesz[,3])),"0and1")
colnames(cellnamesz)<-colnames(cellnamesy)
#Combination
pdmat<-as.matrix(rbind(cellnamesy,cellnamesz))
pddf<-as.data.frame(pdmat)
pddf$pt<-as.numeric(as.character(pddf$pt))
pd<-new("AnnotatedDataFrame",data=pddf)

#Numeric Matrices

#Yang

row.names(loadmaty)<-loadmaty[,1]
maty<-loadmaty[2:length(loadmaty[,1]),match(rownames(cellnamesy),loadmaty[1,])]
colnames(maty)<-rownames(cellnamesy)
maty<-as.matrix(maty)
matchind<-match(rownames(fdmat),rownames(maty))
maty<-maty[matchind,]
nummaty<-mat2numericmat(maty)
#idandname<-function(idnamemat, geneid, genename)
row.names(nummaty)<-idandname(fdmat, rownames(nummaty),"NA")

#Zonation
tidymatz<-zonatedmat2[6:length(zonatedmat2[,1]),]
matchind<-match(rownames(tidymatz),fdmat[,1])
matchind<-matchind[-which(is.na(matchind))]
tidymatz<-tidymatz[matchind,]
row.names(tidymatz)<-rownames(fdmat)[matchind]
matz<-as.matrix(tidymatz[,-1])
nummatz<-mat2numericmat(matz)
row.names(nummatz)<-idandname(fdmat, rownames(nummatz),"NA")


```

#Seurat Implementation

```{r Seurat obj, warning=FALSE, message=FALSE}
#Yang cell dataset
cdsy <- CreateSeuratObject(raw.data = nummaty)
cdsy <- NormalizeData(object = cdsy)
cdsy <- ScaleData(object = cdsy)
cdsy <- FindVariableGenes(object = cdsy, do.plot = FALSE)
#Zonation cell dataset
cdsz <- CreateSeuratObject(raw.data = nummatz)
cdsz <- NormalizeData(object = cdsz)
cdsz <- ScaleData(object = cdsz)
cdsz <- FindVariableGenes(object = cdsz, do.plot = FALSE)

# we will take the union of the top 2k variable genes in each dataset for
# alignment note that we use 1k genes in the manuscript examples, you can
# try this here with negligible changes to the overall results
hvg.cdsy <- rownames(x = head(x = cdsy@hvg.info, n = 4437))
hvg.cdsz <- rownames(x = head(x = cdsz@hvg.info, n = 4437))
hvg.union <- union(x = hvg.cdsy, y = hvg.cdsz)

cdsz@meta.data[, "protocol"] <- "Zonation"
cdsy@meta.data[, "protocol"] <- "Yang"

```


#Running CCA

```{r CCA, fig.height=6, fig.width=12}
cdscomb<- RunCCA(object = cdsy, object2 = cdsz, genes.use = hvg.union)
p1 <- DimPlot(object = cdscomb, reduction.use = "cca", group.by = "protocol", pt.size = 0.5, 
    do.return = TRUE)
p2 <- VlnPlot(object = cdscomb, features.plot = "CC1", group.by = "protocol", do.return = TRUE)
plot_grid(p1, p2)

PrintDim(object =cdscomb, reduction.type = "cca", dims.print = 1:2, genes.print = 10)
```

Choosing which CCs to use 

```{r cc choice, message=FALSE,warning=FALSE}
p3 <- MetageneBicorPlot(cdscomb, grouping.var = "protocol", dims.eval = 1:20, 
    display.progress = FALSE)
```

20 CCs are appropriate to use

Lets look at heatmaps for first 9 CCs

```{r heatmap, fig.height=12, fig.width=12}
DimHeatmap(object = cdscomb, reduction.type = "cca", 
    dim.use = 1:9, do.balanced = TRUE)

```


```{r, message=FALSE, warning=FALSE}
cdscomb <- AlignSubspace(cdscomb, reduction.type = "cca", grouping.var = "protocol", 
    dims.align = 1:20)
```


Violin plots to verify alignment:
```{r vln plots}
p1 <- VlnPlot(object = cdscomb, features.plot = "ACC1", group.by = "protocol", 
    do.return = TRUE)
p2 <- VlnPlot(object = cdscomb, features.plot = "ACC2", group.by = "protocol", 
    do.return = TRUE)
plot_grid(p1, p2)
```

The Yang data still seems a little skewed, but appears to be better aligned than the 1st.

Now a tSNE is ran on the combined dataset

```{r tSNE}
cdscomb<-RunPCA(cdscomb)
PCAPlot(cdscomb)
cdscomb <- RunTSNE(cdscomb, reduction.use = "cca.aligned", dims.use = 1:20, 
    do.fast = T)
```

```{r tSNE plot}
p1 <- TSNEPlot(cdscomb, do.return = T, pt.size = 0.5, group.by = "protocol")
print(p1)
cdscomb <- FindClusters(cdscomb, reduction.type = "cca.aligned", 
    resolution = .7, dims.use = 1:20)
p2 <- TSNEPlot(cdscomb, do.label = T, do.return = T, pt.size = 0.5)
plot_grid(p1, p2)
```

We see that there are 4 clusters found, and we look at markers that identify hepatic expression.


```{r tSNE1, fig.height=12, fig.width=12}
clusterb<-c("Tbx3", "Id3",  "Etv5", "Lgr5")#hepatoblast genes
clusterc<-c("Cps1", "Ppara", "Apoh", "Cyp2d10", "Cyp2d26")#cluster c genes

FeaturePlot(object = cdscomb, features.plot=c(clusterb,clusterc) )

```

We see that cluster 3 can be identified as hepatocytes

Conserved markers are identified and plotted in a similar fashion, we use cluster 3 as an example.

```{r}
nk.markers <- FindConservedMarkers(cdscomb, ident.1=3, grouping.var = "protocol", 
    print.bar = FALSE)
ccamarkers<-rownames(nk.markers)
FeaturePlot(object = cdscomb, features.plot=ccamarkers[1:12] )
```

```{r}
write.table(ccamarkers, "ccamarkers.txt")
tsnevals<-as.matrix(cdscomb@dr$tsne@cell.embeddings)
write.xlsx(tsnevals, "tSNEvals2.xlsx")
ccavals<-as.matrix(cdscomb@dr$cca@cell.embeddings)
ccavals<-ccavals[,1:2]
write.xlsx(ccavals,"ccavals.xlsx")
ccamarkers<-as.matrix(cdscomb@dr$cca@gene.loadings.full)
ccamarkers<-ccamarkers[,1:2]
write.xlsx(ccamarkers, "ccamarkers.xlsx")
accamarkers<-as.matrix(cdscomb@dr$cca.aligned@cell.embeddings)
accamarkers<-accamarkers[,1:2]
write.xlsx(accamarkers, "accamarkers.xlsx")
```